---
title: "Final_quiz_predictionCodes"
author: "wenting"
date: "May 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Study Goal
  is to predict the manner in which the participants did the exercise by using data from accelerometers on the belt, forearm, arm, and dumbell. 
This is the "classe" variable in the training set.
  
### Step 1  read data
```{r}
setwd("/Users/Final_Assignment")
training=read.csv("pml-training.csv",sep=",",head=T,stringsAsFactors = F,na.strings=c("","NA"))
testing=read.csv("pml-testing.csv",sep=",",head=T,stringsAsFactors = F,na.strings=c("","NA"))
dim(training)
dim(testing)
```

### Step 2  clean data
#### 2.1 remove variables with mostly missing values
```{r}
mostly_data<-apply(!is.na(training),2,sum)>19621
training<-training[,mostly_data]
testing<-testing[,mostly_data]
```

#### 2.2 remove column that is simply an index, timestamp
```{r}
training=training[,c(2,8:60)]
testing=testing[,c(2,8:60)]
dim(training)
```

#### 2.3 change variables to numeric and factor
```{r,dependson="createData",fig.height=5,fig.width=5}
for (i in 2:53){
	training[,i]=as.numeric(as.character(training[,i]))
}
training$classe=as.factor(as.character(training$classe))
training$user_name=as.factor(as.character(training$user_name))
```

### 3 - Training Models - RandomForest
#### Setting the seed for reproducibility
```{r}
set.seed(1235)
```

#### load packages 
```{r}
library(caret)
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
```

#### 3.1 develop training model with 5-fold Cross-Validation
```{r}
fitControl <- trainControl(method = "cv",
                           number = 5,
                           allowParallel = TRUE)

fit <- train(classe~., data=training,method="rf",trControl = fitControl,prox=TRUE)
```

#### stop cluster
```{r}
stopCluster(cluster)
```

#### 3.2 get model
```{r}
fit
fit$resample
fit$finalModel
```

#### 3.3 evaluate the final accuracy in cross validation
```{r}
confusionMatrix.train(fit)
plot(fit,main="Accuracy by Predictor Count")
varImpPlot(fit$finalModel, main="Variable Importance Plot: Random Forest")
```

### 4.  Predicting classe of the 20 Test Cases

#### 4.1 prepare new test cleaned data
```{r}

for (i in 2:53){
	testing[,i]=as.numeric(as.character(testing[,i]))
}

testing$user_name=as.factor(testing$user_name)
```

#### 4.2 Predicting new values

```{r predForest, dependson="centers",fig.height=4,fig.width=4,cache=TRUE}
pred <- predict(fit,testing)
pred
```
#### Summary: the rf model was applied to predict/classy the final 20 test samples in quiz and it is 100% accuarate.


